---
title: "Cumulative-Killed-Shot Graphics 2014-2022"
author: "MCC"
date: "`r Sys.Date()`"
output: pdf_document
---

## Setup

```{r message=FALSE, warning=FALSE, include=FALSE}
library(readr)
library(lubridate)
library(ggplot2)
library(tidyverse)
library(knitr)
library(here)

here::i_am("Desktop/gun_violence_research")

pagebreak <- function() {
  if(knitr::is_latex_output())
    return("\\newpage")
  else
    return('<div style="page-break-before: always;" />')
}
```

## Read Gun Violence dataframes

**read_gva_data**

```{r}
read_gva_data <- function(i) {
  file_year = paste("~/Desktop/gun_violence_research/001_data/001A_ORIGINAL/gva-", i, ".csv", sep="")
  # print(file_year)
  gva_df <- read_csv(file_year, 
        col_types = cols(`Incident ID` = col_skip(), 
        `Incident Date` = col_date(format = "%B %d, %Y"), 
        State = col_skip(), `City Or County` = col_skip(), 
        Address = col_skip(), Operations = col_skip()))
  
  # Simplify Names & Order data frame by 'ID' (incident date)
  names(gva_df) <- c("ID", "Killed", "Injured")
  gva_df <- gva_df[order(gva_df$`ID`), ]
  
  ## Calculate Killed/Injured/Grand Totals
  date <- ymd(gva_df$`ID`)
  gva_df$Days <- yday(date) - 1 # so Jan 1 = day 0 
  gva_df$Cum_Killed <- cumsum(gva_df$Killed)
  gva_df$Cum_Injured <- cumsum(gva_df$Injured)
  gva_df$Grand_Total <- gva_df$Cum_Killed + gva_df$Cum_Injured
  return(gva_df)
}
```

`r pagebreak()`

## Graph of Cumulative Deaths vs Days for Given Year

**graph_deaths**

```{r message=FALSE, warning=FALSE}
graph_deaths <- function(gva_df, subject, i){
  #print("graph_deaths")
  y_col <- paste('gva_df$', subject, sep="")
  title <- paste(i, ": Cumulative ", subject, "v. Day Per Year", sep="")
  y_title <- paste("Cumulative ", subject, sep="")
  require(ggplot2)
  plot(ggplot(gva_df, aes(x=gva_df$Days, y=y_col)) +
       geom_line() +
       ggtitle(title) +
       labs(x='Day No. of 365 Days', y=y_title) +
       geom_smooth(method = "lm"))
}
```

## Regression Numbers For Killed vs Days for Given Year

**calc_regression_killed**

```{r}
calc_regression_killed <- function(gva_df, i){
  #print("calc_regression_killed")
  #print(names(gva_df))
  #fit simple linear regression model
  model <- lm(gva_df$Cum_Killed ~ gva_df$Days)
  print(paste("Regression Values For Year: ", i, sep=""))
  print(model)
}
```

## Graph Total Shot (Killed + Injured) vs Days For Given Year

**graph_total_shot**

```{r message=FALSE, warning=FALSE}
graph_total_shot <- function(gva_df, i){
  #print("graph_total_shot")
  require(ggplot2)
  plot(ggplot(gva_df, aes(x=gva_df$Days, y=gva_df$Grand_Total)) +
       geom_line() +
       ggtitle("Cumulative Shot (Killed + Injured) Per Year ",
               subtitle = i) +
       labs(x="Day No. of 365 Days", y="Cumulative Total Shot (Killed + Injured)") +
       geom_smooth(method = "lm"))
}
```

## Calculate Regression Numbers For Total Shot vs Days for Given Year

**calc_regression_total_shot**

```{r}
calc_regression_total_shot <- function(gva_df, i){
  #print("calc_regression_total_shot")
  #fit simple linear regression model
  model <- lm(gva_df$Grand_Total ~ gva_df$Days)
  print(paste("Regression Values For Year: ", i, sep=""))
  print(model)
}
```
 


## Main
```{r message=FALSE, warning=FALSE}
for (i in 2014:2021) {
  gva_df <- data.frame()
  gva_df <- read_gva_data(i)
  subject = "Killed"
  pagebreak()
  graph_deaths(gva_df, subject, i)
  calc_regression_killed(gva_df, i)
  graph_total_shot(gva_df, i)
  calc_regression_total_shot(gva_df, i)
}
```

| YEAR | Killed    | Total    | AVG |
| :--- | :-------- | :------- | :-- |
| 2014 | 0.7628257 | 4.024422 | 5.3 |
| 2015 | 1.059014  | 5.157974 | 4.9 |
| 2016 | 1.384732  | 6.204191 | 4.5 |
| 2017 | 1.232796  | 6.456688 | 5.2 |
| 2018 | 1.062215  | 5.377669 | 5.1 |
| 2019 | 1.348324  | 6.540035 | 4.9 |
| 2020 | 1.588776  | 10.33217 | 6.5 |
| 2021 | 56.16621  | 151.9159 | 2.7 |
|      |           | Avg      | 4.9 |













