---
title: "Mother_Jones_Gun_Violence_DB"
author: "MCC"
date: "`r Sys.Date()`"
output: html_document
---


```{r}
Mother_Jones_Gun_Violence_DB <- read.csv("~/Dropbox/1computer_projects/DATA/Gun_violence_data/Mother Jones - Mass Shootings Database, 1982 - 2021 - Sheet1.csv")

#View(`Mother.Jones...Mass.Shootings.Database,.1982...2021...Sheet1`)

str(Mother_Jones_Gun_Violence_DB, give.attr = FALSE)
```

There are 125 observations with 24 features.

Only six of the features are numeric or integer.     
1. fatalities    
2. injured    
3. total victims    
4. Latitude     
5. Longitude    
6. Year    

The remaining 18 are character vectors.   

```{r}
dim(Mother_Jones_Gun_Violence_DB)
```


### Let's carry out a six-number summary of the six numeric features

```{r}
mj_db_numeric <- Mother_Jones_Gun_Violence_DB[c(5,6,7)]

summary(mj_db_numeric)
```

- Of the three features, there is one outlier. This is the Las Vegas Shooting which had 58 fatalities, 546 people injured which is a total of 604 people. 

- If we consider the 'Deadliest Terror Attacks' table below for comparison. We find that the 58 fatalities from the Las Vegas shooting would be the tenth most fatal incident.
OR
- If we use the total victims we see that 604 victims would rank **fifth**.

```{r}
colSums(mj_db_numeric)
```

See: https://www.worldatlas.com/articles/the-deadliest-terror-attacks-in-the-united-states.html

The Deadliest Terror Attacks In The United States

Rank | Name Of Incident | Date | Fatalities | Injuries | Location(s)   
-----|------------------|------|------------|----------|------------    
1 | September 11 attacks | September 11, 2001 | 2,996 (inc. 19 suspects) | 6,000+ | New York City, New York, Arlington, Virginia, Shanksville, Pennsylvania   
2 | Lawrence massacre | August 21, 1863 | 204 (inc. 40 suspects) | 1 | Douglas County, Kansas  
3 | Oklahoma City bombing | April 19, 1995 | 168 | 680+ | Oklahoma City, Oklahoma  
4 | Tulsa race riot | June 1, 1921 | 150–300 | 800+ | Greenwood, Tulsa, Oklahoma   
5 | Mountain Meadows massacre | September 11, 1857 | 120+ | 17+ | Mountain Meadows, Utah Territory  
6 | Colfax massacre | April 13, 1873 | 108 (inc. 3 suspects) | unknown | Colfax, Louisiana  
7 | Gnadenhutten massacre | March 8, 1782 | 96 | 2 | Gnadenhutten, Ohio  
8 | Wiyot massacre | February 26, 1860 | 80+ | unknown | Indian Island (Humboldt Bay), California  
9 | Hillabee massacre | November 13, 1813 | 64 | 29 | Hillabee, Alabama  
10 | Orlando nightclub shooting | June 12, 2016 | 50 (inc. 1 suspect) | 53 | Orlando, Florida  
11 | Bath School disaster | May 18, 1927 | 45 (inc. 1 suspect) | 58 | Bath Township, Michigan  
12 | Asbill massacre | May 15, 1854 | 40 | unknown | Mendocino County, California   
13 | Wall Street bombing | September 16, 1920 | 38 | 143 | New York City, New York  
14 | Pigeon Roost massacre | September 3, 1812 | 24 | unknown | Pigeon Roost Village, Northwest Territory  
15 | Wilmington insurrection of 1898 | November 10, 1898 | 22+ | unknown | Wilmington, North Carolina  
16 | Los Angeles Times bombing | October 1, 1910 | 21 | 100+ | Los Angeles, California  
17 | San Bernardino attack | December 2, 2015 | 16 (inc. 2 suspects) | 24 | San Bernardino, California  
18 | Fort Hood shooting | November 5, 2009 | 13 | 33 (inc. 1 suspect) | Fort Hood, Texas  
19 | Big Bottom massacre | January 2, 1791 | 12 | several | near Stockport, Ohio  
20 | Haymarket affair | May 4, 1886 | 11 | 130+ | Chicago, Illinois  

- To Do
1. What are the largest terrorist attacks or mass casualities in the US.   
   - OK City, Federal Building, ?? numbers    
   - World trade airline attack ?? numbers    
   - ARE THERE ANY OTHERS?    
   
FYI: https://www.history.com/news/deadliest-events-united-states   

| Rank | Event                                      | Fatalities |
| :--- | :----------------------------------------- | :--------- |
| 1.   | The COVID-19 Pandemic                      | 950,000    |
| 2.   | The US Civil War                           | 750,000    |
| 3.   | The HIV/AIDS Epidemic                      | 700,000    |
| 4.   | The 1918 Flu Pandemic                      | 675,000    |
| 5.   | World War II                               | 405,400    |
| 6.   | World War I                                | 116,516    |
| 7.   | The Vietnam War                            | 58,220     |
| 8.   | The Korean War                             | 36,914     |
| 9.   | The 1900 Galveston Hurricane               | 8,000      |
| 10.  | The 1906 San Francisco Earthquake and Fire | 3,000      |
 

### Produce Histograms of Fatalities, Injured, and Total Victims

```{r}
hist(mj_db_numeric[,"fatalities"])
```

```{r}
hist(mj_db_numeric[,"injured"])
```

```{r}
hist(mj_db_numeric[,"total_victims"])
```



```{r}
library(readr)
MJ_Short_Database <- read_csv("Mother Jones - Mass Shootings Database, 1982 - 2021 - Sheet1.csv", 
    col_types = cols(summary = col_skip(), 
        age_of_shooter = col_character(), 
        prior_signs_mental_health_issues = col_skip(), 
        mental_health_details = col_skip(), 
        weapons_obtained_legally = col_skip(), 
        where_obtained = col_skip(), sources = col_skip(), 
        mental_health_sources = col_skip(), 
        sources_additional_age = col_skip(), 
        latitude = col_skip(), longitude = col_skip()))

#View(Mother_Jones_Mass_Shootings_Database_1982_2021_Sheet1)
#

str(MJ_Short_Database)
```















